---
title: "Analysis of single cell ferret pancreatic ductal differentiation"
author: "Kristen Wells"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    highlight: "tango"
    df_print: "paged"
    code_folding: "hide"
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = F, 
  warning = F,
  comment = ""
)
library(here)

knitr::opts_knit$set(root.dir = here())
```

# Summary

```{r, include = FALSE}
library(Seurat)
library(tidyverse)
library(cowplot)
library(here)
library(scAnalysisR)
library(knitr)
library(LaCroixColoR)

# Set theme
ggplot2::theme_set(ggplot2::theme_classic(base_size = 10))

# Sample information that also includes if it is combined or not
samples <- list("WT_D2" = c("single"),
                "WT_D5" = c("single"),
                "WT_D7" = c("single"),
                "WT_D9" = c("single"),
                "WT_D14" = c("single"),
                "CFKO_D2" = c("single"),
                "CFKO_D5" = c("single"),
                "CFKO_D7" = c("single"),
                "CFKO_D9" = c("single"),
                "CFKO_D14" = c("single"))

ADT <- FALSE

# Set directories
base_dir <- here()

# Change to be the cell types of the analysis
# colors from https://medialab.github.io/iwanthue/
byrnes_colors <- c("Acinar" = "#D4405B",
                   "Fev_hi" = "#75B140",
                   "Alpha" = "#885FCF",
                   "Ductal" = "#A5903E",
                   "Ngn3_pos" = "#C65CAC",
                   "Prolif_acinar" = "#55A470",
                   "Prolif_ductal" = "#767FC9",
                   "undetermined" = "#D3D3D3")

baron_colors <- c("ductal" = "#A5903E",
                  "acinar" = "#D4405B")

tabula_muris_colors <- c("pancreatic_acinar_cell" = "#D4405B",
                         "pancreatic_stellate_cell" = "#CC6D3A",
                         "pancreatic_ductal_cell" = "#A5903E",
                         "endothelial_cell" = "#45B0CF",
                         "undetermined" = "#D3D3D3")

qadir_colors <- c("immune_cells" = "#D3D3D3",
                  "progenitor_like_cells" = "#297878",
                  "transitional_to_acinar1" = "#874652",
                  "transitional_to_acinar2" = "#CC1B3B",
                  "centroacinar" = "#78295D",
                  "Small_ducts" = "#B37422")

muraro_colors <- c("ductal" = "#A5903E",
                  "acinar" = "#D4405B",
                  "undetermined" = "#D3D3D3")

all_colors <- c("Acinar" = "#D4405B",
                "Ductal" = "#A5903E",
                "ductal" = "#A5903E",
                "Prolif_acinar" = "#55A470",
                "Prolif_ductal" = "#767FC9",
                "progenitor_like_cells" = "#297878",
                "transitional_to_acinar1" = "#874652",
                "transitional_to_acinar2" = "#CC1B3B",
                "centroacinar" = "#78295D")

# Change to be the samples of the analysis
sample_colors <- LaCroixColoR::lacroix_palette("Coconut", 10)
names(sample_colors) <- c("WT_D2", "WT_D5", "WT_D7", "WT_D9", "WT_D14",
                          "CFKO_D2", "CFKO_D5", "CFKO_D7", "CFKO_D9", "CFKO_D14")

# Read in all data
sample_list <- lapply(names(samples), function(x){
  print(x)
  base_dir_proj <- file.path(base_dir, "results", x)
  save_dir <- file.path(base_dir_proj, "R_analysis")
  seurat_data <- readRDS(file.path(save_dir, "rda_obj", "seurat_processed.rds"))
  unprocessed_path <- file.path(save_dir, "rda_obj", "seurat_unfilt.rds")
  doublet_path <- file.path(save_dir, "rda_obj", "seurat_doublet.rds")
  
  nColors <- length(unique(seurat_data$cluster_celltype))
  cluster_colors <- grDevices::colorRampPalette(
    RColorBrewer::brewer.pal(9, "Set1"))(nColors)
   names(cluster_colors) <- unique(seurat_data$cluster_celltype)

  return_data <- list("seurat_object" = seurat_data,
                      "seurat_unprocessed_path" = unprocessed_path,
                      "seurat_doublet_path" = doublet_path,
                      "type" = samples[[x]][1],
                      "cluster_colors" = cluster_colors,
                      "save_dir" = save_dir)
  
})

names(sample_list) <- names(samples)

```

## Quality control {.tabset}
I first check the quality of all samples by looking at the number of reads per cell, the number of features (genes) per cell, and the percent mitochondrial genes per cell. In cases where the cells were dying, we generally see a higher percent mitochondria. I like to see that most cells have less than 10% of reads mapping to mitochondrial reads. The mitochondrial percent was pretty high on most of your samples, so I set the cutoff at 20%

Here it seems that most cells look pretty good. You have ~4,000-5,000 genes per cell. One thing that is a bit strange is that all samples have a second set of cells with around 2,000 genes/cell. This is high enough that I don't want to filter it out in case they are interesting (for example, T cells sometimes have fewer genes/cell than other cells), but in all samples I do see a low quality cluster (look at the UMAPs below), so we may consider filtering these out.

For the moment, all filters are set at the same levels because I wanted to be as consistent as possible with these samples. We may need to go back and change that though.

My filters:
* Percent mitocondria less than 20%
* Number of genes (nFeatureRNA) between 1000 and 8000

```{r "plot-quality", echo = F}
if(ADT){
  quality_template <- file.path(here(),
                              "src/scripts/templates/quality_template_ADT.Rmd")
} else{
quality_template <- file.path(here(),
                              "src/scripts/templates/quality_template.Rmd")  
}

quality_chunks <- names(sample_list) %>%
  map(~knit_expand(file = quality_template, sample = .x))
```

`r knit_child(text = quality_chunks)`

## Doublet Removal {.tabset}
I next identify likely doublets using `DoubletFinder` which attempts to model what doublets would look like based on a mixing of the different clusters. The doublets identified are shown below. I filter these out before continuing the analysis.

```{r "plot-doublet", echo = F}

doublet_template <- file.path(here(),
                              "src/scripts/templates/doublet_template.Rmd")

sample_doublet <- samples[samples == "single"]

doublet_chunks <- names(sample_doublet) %>%
  map(~knit_expand(file = doublet_template, sample = .x))
```

`r knit_child(text = doublet_chunks)`

## PCA {.tabset}
I next do an initial dimensional reduction with using PCA on the top 2000 variable genes. The PCA is not too informative, but it's worth looking at a few metrics.

A) The sample - this will be more interesting once I combine the samples.
B) The number of UMIs seen in each cell. There is some correlation between this and PC1 so we will need to make sure this isn't captured when we make a UMAP.
C) The number of genes seen in each cell. Again, there is some correlation between this and PC1 so we will need to make sure this isn't captured when we make a UMAP.
D) The percent mitochondrial reads per cell. We don't want to see much correlation between this and either of the PCs.
```{r "plot-pca", echo = F}

pca_template <- file.path(here(),
                          "src/scripts/templates/pca_template.Rmd")

pca_chunks <- names(sample_list) %>%
  map(~knit_expand(file = pca_template, sample = .x))
```

`r knit_child(text = pca_chunks)`

## UMAP {.tabset}

I next follow the PCA with a UMAP dimensional reduction. Rather than use genes for UMAP, we use the top PCs (~30 for all of these samples). I'll plot the same metrics for UMAP as for the PCA.

A) The sample - again, not informative until we combine the samples
B) The number of UMIs seen in each cell. There is some correlation between this and the UMAP. It could be technical or cell type specific. We can decided this based on the gene expression patterns.
C) The number of genes seen in each cell. Again, there is some correlation between this and the UMAP. It could be technical or cell type specific. We can decided this based on the gene expression patterns.
D) The percent mitochondrial reads per cell. There isn't much correlation here which is a good sign.


```{r "plot-umap", echo = F}

umap_template <- file.path(here(),
                           "src/scripts/templates/umap_template.Rmd")

umap_chunks <- names(sample_list) %>%
  map(~knit_expand(file = umap_template, sample = .x))
```

`r knit_child(text = umap_chunks)`

## Celltypes {.tabset}
I named celltypes based using `clustifyr` which uses a reference dataset to name clusters. First, I used the mouse orthologs to rename all gene names. Then, average expression was found for each cluster in the reference dataset, then a correlation was run between the reference cluster and our clusters. I used three different references and then combined them based on the top correlation for each cluster.

```{r "plot-celltype", echo = F}

celltype_template <- file.path(here(),
                           "src/scripts/templates/celltype_template.Rmd")

celltype_chunks <- names(sample_list) %>%
  map(~knit_expand(file = celltype_template, sample = .x))
```

`r knit_child(text = celltype_chunks)`
